---
title: "Day3Comp"
format: html
editor: visual
---

## Conditional Statement

```{r}
temp <- 98.2

if (temp > 97) {
  print("its real hot, jump in the ocean")
} else
  print("no temperature warning")
```

Create an if else statement for blue whale lengths.

create a variable blue whale, which has a value of 40 meters.

add an if else statement that returns "GIANT" if \>35

medium if 20\<x\<35 or

small if \<=20

```{r}
whale_length <- 26

if (whale_length >35) {
  print("GIANT")
} else if (whale_length >= 20 & whale_length<=35) {
  print("MEDIUM")
} else if (whale_length <= 20) {print("SMALL")}

```

## Switch Statement

```{r}
housing_type <- "single"

switch(housing_type,
       "single" = ("400"),
       "double" = ("750"),
       "whole" = ("1100"))

if (housing_type == "whole") {
  print("1100")
} else if (housing_type == "double") {
  print("750")
}else if (housing_type == "single") {
  print("400")
}
```

## For Loops

```{r}
height <- c(58, 61, 65, 72, 89, 94, 101)

for (i in seq_along(height)) {
  half_height = height[i] / 2
  print(half_height)
}

## for each value in the height vector, return the sum of the value + the one right after it

height <- c(58, 61, 65, 72, 89, 94, 101)

for (i in seq_along(height)) {
  sum_height = height[i]+height[i+1]
  print(sum_height)
}

for (i in seq(1:length(height)-1)) {
  sum_height = height[i]+height[i+1]
  print(sum_height)
}

## write a for loop that calculates the value of height for each element divided by the mean of the entire height vector

height <- c(58, 61, 65, 72, 89, 94, 101)

for (i in seq_along(height)) {
  standardized_ht = height[i]/mean(height)
  print(standardized_ht)
}
```
